---
title: "Reproducible_Document"
author: "rostools"
format: html
---

Special "setup" code chunk: When you want your Quarto to run one specific, important code chunk before everything else - like loading in tidyverse
```{r setup}
#load in tidyverse; is not in itself a package but rather a meta-package
library(tidyverse)

#for data separated by semicolons, we use the read_csv2() function instead of the read_csv()
post_meal_data <- read_csv2(here::here("data/post-meal-insulin.csv"))
```

## About me

- Cecilie Kynding
- Novo Nordisk Foundation Center for Basic Metabolic Research (CBMR), University of Copenhagen

I'm a pre-PhD, doing a maternity cover in the Rodent Metabolic Phenotyping Platform at **CBMR**. Later this year, I'll start my PhD project, investigating *the potential of a growth hormone and semaglutide combination treatment regimen*.

## Simple code

```{r}
3*3
```
Man kan bruge gitignore til at skjule personfølsomme data eller data, der ikke skal uploades til Git. Man behøver f.eks. ikke at uploade html eller word-filer, da disse genereres ud fra vores Quarto-filer. Så længe vi har vores originale Quarto-filer, kan vi altid regenerere html eller word-filer senere.

Dette gøres ved at indtaste dette i konsollen: 
usethis::use_git_ignore(c("*.html", "*_files/"))

## Testing for Git

The weather is insanely cold and windy, yet it is a sunny and beautiful day.

## Session 17 - Importing data into R

For downloading data, we want to type it in the console rather than in a code chunk in our document, since if we run it in a code chunk in our document, we will download the file every time we render the document, instead of just downloading it once in the console.

## Showing the data

```{r}
post_meal_data

glimpse(post_meal_data)
```


