---
title: "Reproducible_Document"
author: "rostools"
format: html
---

Special "setup" code chunk: When you want your Quarto to run one specific, important code chunk before everything else - like loading in tidyverse
```{r setup}
#load in tidyverse; is not in itself a package but rather a meta-package
library(tidyverse)
```

## About me

- Cecilie Kynding
- Novo Nordisk Foundation Center for Basic Metabolic Research (CBMR), University of Copenhagen

I'm a pre-PhD, doing a maternity cover in the Rodent Metabolic Phenotyping Platform at **CBMR**. Later this year, I'll start my PhD project, investigating *the potential of a growth hormone and semaglutide combination treatment regimen*.

## Simple code

```{r}
3*3
```
Man kan bruge gitignore til at skjule personfølsomme data eller data, der ikke skal uploades til Git. Man behøver f.eks. ikke at uploade html eller word-filer, da disse genereres ud fra vores Quarto-filer. Så længe vi har vores originale Quarto-filer, kan vi altid regenerere html eller word-filer senere.

Dette gøres ved at indtaste dette i konsollen: 
usethis::use_git_ignore(c("*.html", "*_files/"))

## Testing for Git

The weather is insanely cold and windy, yet it is a sunny and beautiful day.
