---
title: "Reproducible_Document"
author: "rostools"
format: html
---

Special "setup" code chunk: When you want your Quarto to run one specific, important code chunk before everything else - like loading in tidyverse
```{r setup}
#| message: false
#| warning: false

# load in tidyverse; is not in itself a package but rather a meta-package
library(tidyverse)

# for data separated by semicolons, we use the read_csv2() function instead of the read_csv()
post_meal_data <- read_csv2(here::here("data/post-meal-insulin.csv"))

# when splitting up data like this, always choose the id column first to make sure it's not lost which id each information relates to
small_post_meal_data <- select(post_meal_data, id, Age, BMI, Group, auc_pg, auc_ins, glykemi)

tidier_post_meal_data <- small_post_meal_data |>
  rename(
    auc_glucose = auc_pg,
    auc_insulin = auc_ins
  ) |>
  filter(glykemi == 0) |>
  select(-glykemi)
```

## About me

- Cecilie Kynding
- Novo Nordisk Foundation Center for Basic Metabolic Research (CBMR), University of Copenhagen

I'm a pre-PhD, doing a maternity cover in the Rodent Metabolic Phenotyping Platform at **CBMR**. Later this year, I'll start my PhD project, investigating *the potential of a growth hormone and semaglutide combination treatment regimen*.

## Simple code

```{r}
3 * 3
```
Man kan bruge gitignore til at skjule personfølsomme data eller data, der ikke skal uploades til Git. Man behøver f.eks. ikke at uploade html eller word-filer, da disse genereres ud fra vores Quarto-filer. Så længe vi har vores originale Quarto-filer, kan vi altid regenerere html eller word-filer senere.

Dette gøres ved at indtaste dette i konsollen: 
usethis::use_git_ignore(c("*.html", "*_files/"))

## Testing for Git

The weather is insanely cold and windy, yet it is a sunny and beautiful day.

## Session 17 - Importing data into R

For downloading data, we want to type it in the console rather than in a code chunk in our document, since if we run it in a code chunk in our document, we will download the file every time we render the document, instead of just downloading it once in the console.

## Showing the data

```{r}
post_meal_data

glimpse(post_meal_data)
```

## Plot one continuous variable

```{r}
#| fig-cap: "Distribution of BMI"
#| label: fig-bmi-histo
ggplot(post_meal_data, aes(x = BMI)) +
  geom_histogram()
```
Reference to figure, using added label in code chunk
See @fig-bmi-histo

## Exercise: discrete plots

```{r}
#| fig-cap: "Group"
#| label: fig-group-bar
ggplot(post_meal_data, aes(x = Group)) +
  geom_bar()
```

See @fig-group-bar above for a cool plot!

## Plotting two discrete variables

```{r}
ggplot(post_meal_data, aes(x = Group, fill = as.character(glykemi))) +
  geom_bar(position = position_dodge())

# as.character needs to be added in front of glykemi in the 'fill = ', since glykemi is a numerical information
```
## Side by side plots

```{r fig-bmi-glycemia}
#| fig-cap: "BMI and glycemia, side by side"
#| fig-subcap:
#| - "Distribution of BMI."
#| - "Number of those with glycemia"
#| layout-ncol: 2
ggplot(post_meal_data, aes(x = BMI)) +
  geom_histogram()
ggplot(post_meal_data, aes(x = glykemi)) +
  geom_bar()
```

## Plotting two continuous variables

```{r}
#| fig-cap: "Scatterplot of BMI and the area under the curve for glucose"
#| label: fig-bmi-auc-pg
ggplot(post_meal_data, aes(x = BMI, y = auc_pg)) +
  geom_point() +
  geom_smooth()
```

## Selecting data

```{r}
select(post_meal_data, Age)

select(post_meal_data, Age, BMI, Weight)

select(post_meal_data, -Group)
```

## Select helpers

```{r}
# use "" when referring to something that's not a number
select(post_meal_data, starts_with("PG"))

select(post_meal_data, ends_with("DXA"))

# even though we're working with a number (0), R considers it a word/character because it's part of a column name
select(post_meal_data, contains("0"))
```

## Renaming columns

```{r}
# pipe shortcut: command + shift + M
small_post_meal_data |>
  rename(
    auc_glucose = auc_pg,
    auc_insulin = auc_ins
  )
```

## Select and rename tasks

```{r}
# Task 1.
post_meal_data |>
  select(contains("OGTT"))
# 13 columns

# Task 2.
post_meal_data |>
  select(contains(".."), -contains("..."))

# Task 3.
post_meal_data |>
  rename(
    P.Glucose.minus5.OGTT = P.Glucose..5.OGTT,
    Insulin.minus5.OGTT.X = Insulin..5.OGTT.X
  )

# Task 4.
basic_info <- select(post_meal_data, id, BMI, Length, Weight, Age)
rename(basic_info, Height = Length)

post_meal_data |>
  select(id, BMI, Length, Weight, Age) |>
  rename(Height = Length)
```

## Filtering data

```{r}
small_post_meal_data |>
  filter(glykemi == 1)

small_post_meal_data |>
  filter(BMI >= 25)

small_post_meal_data |>
  filter(Group == "CTR")

# using | for either/or, and & (or ,) for both arguments to be true
small_post_meal_data |>
  filter(Group == "CTR" | BMI >= 25)
```
